### 2020-06-05 README.Rmd
include type in tablename to avoid ambiguity

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(neonstore)

@@

The goal of neonstore is to provide quick access and persistent storage of NEON data tables.

@@

```{r example, cache=TRUE}
library(neonstore)
neon_download("DP1.10003.001")

@@

Now, view your store of NEON products.  These will persist between sessions,
so you only need to download once, or to retrieve updates.
(Note: the individual files making up a data product can all be listed by
`neon_index()`).

@@

Read in all the component tables into a single data.frame

```{r}
neon_read("brd_countdata")
```

@@

Because `neonstore` only stores raw data products as returned from the NEON API,
this makes it easy to track provenance and compare results across other analyses
of the NEON data.

@@

>>>>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  cache = TRUE
)
library(neonstore)
Sys.setenv("NEONSTORE_HOME" = tempfile())
```

@@

`neonstore` provides quick access and persistent storage of NEON data tables.

@@

```{r}
library(neonstore)
neon_download("DP1.10003.001")

@@

Now, view your store of NEON products:

@@

These will persist between sessions, so you only need to download once
or to retrieve updates. `neon_store()` can take arguments to filter by product
or pattern in table name, e.g. `neon_store(table = "brd")`.

Once you determine the table of interest, you can
read in all the component tables into a single `data.frame`

```{r}
neon_read("brd_countdata-expanded")
```

@@

## Provenance

Because `neonstore` only stores raw data products as returned from the NEON API,
it can easily determine which files have already been downloaded, and only
download new files without requiring the user to specify specific dates. 
(It must still query the API for all the metadata in the requested date range).
This same modular approach also makes it easy to track _data provenance_, an 
essential element of reproduciblity in comparing results across other analyses
of the NEON data.

We can list precisely which component files are being read in by `neon_read()`
by consulting `neon_index()`:

```{r}
raw_files <- neon_index(table = "brd_countdata-expanded")
raw_files
```

`neon_read()` is a relatively trivial function that simply passes this file 
list to `vroom::vroom()`, a fast, vectorized parser that can easily read in 
a single table that is broken into many separate files.

Imagine instead that we use the common pattern of downloading
these raw files, stacks and possibly cleans the data, saving only this derived
product while discarding the individual files.  Now imagine a second researcher,
at some later date, queries the API over the same reported range of dates and 
sites, uses the same software package to stack the tables, only to discover the
resulting table is somehow different from ours (e.g. by comparing file hashes).
Pinpointing the source of the discrepancy would be challenging and 
labor-intensive.

In contrast, the same detective-work would be easy with the `neonstore` file
list.  We can confirm if the API had returned the same number of 
raw files with the same names; and better, can verify integrity of the contents
by comparing hashes of files now being returned to those recorded
by `neon_index()`.  In this way, we could determine if any additional files
had been included or pinpoint any files that may have changed.

As such, users might want to store the `neon_index()` `data.frame` for the
table(s) they have used as part of their analysis, including the individual
file hashes.  One can also generate a zip of all the data files for 
archival purposes. (Note that NEON is an Open Data provider, see
[LICENCE](https://www.neonscience.org/data/about-data/data-policies).)

```{r message=FALSE}
write.csv(raw_files, "index.csv")
zip("brd_countdata.zip", raw_files$path)
```

## Data citation

Generate the appropriate citation for your data:

```{r}
neon_citation()
```

@@

```{r include=FALSE}
unlink("brd_countdata.zip")
unlink("index.csv")
Sys.unsetenv("NEONSTORE_HOME")
```
